---
title: "Springboard Capstone Project"
author: Gautham Sista
output: html_notebook
---

This notebook shows the analysis performed as part of my Springboard Capstone Project for the Introduction to Data Science course. 

1a. Reading in the San Francisco Crime dataset

```{r}
sf <- read.csv("/Users/gsista/Desktop/Springboard/DataScience/Data/sf_crime.csv")
sf$Dates <- as.POSIXct(sf$Dates, format="%Y-%m-%d %H:%M:%S")
str(sf)
```

```{r}
sf
```

1b. Filtering the San Francisco dataset to crimes that happened on and after January 1st, 2013
```{r}
sf1 <- subset(sf, Dates >= "2013-01-01 00:00:00")
sf1$LatLong <- NA
sf1$LatLong <- paste(sf1$Y, sf1$X, sep = ", ")
```

1c. Store the unique list of LatLong in a variable
```{r}
uniqueLatLong <- unique(sf1$LatLong)
uniqueLatLong
```


```{r}
sf$DayOfWeek <- factor(sf$DayOfWeek, levels=c('Monday', 'Tuesday', 'Wednesday', 'Thursday', 'Friday', 'Saturday', 'Sunday'))
library(ggplot2)
ggplot(sf, aes(x = DayOfWeek)) +
  geom_bar() +
  geom_text(stat='count',aes(label=..count..),vjust=-1)
```

```{r}
sf$DayOfWeek <- factor(sf$DayOfWeek, levels=c('Monday', 'Tuesday', 'Wednesday', 'Thursday', 'Friday', 'Saturday', 'Sunday'))
ggplot(sf, aes(x = Category, fill = DayOfWeek)) +
  geom_bar() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```


2. Reading the SF weather dataset

```{r}
weather <- read.csv("/Users/gsista/Desktop/Springboard/DataScience/Springboard-Capstone-Project/Data/weather.csv")
head(weather)
```

3. Reading in Median House Prices by zipcode dataset

```{r}
library(dplyr)
mhp <- read.csv("/Users/gsista/Desktop/Springboard/DataScience/Springboard-Capstone-Project/Data/Zip_MedianListingPrice_AllHomes.csv")
str(mhp)
sfmhp <- filter(mhp, City == 'San Francisco')
sfmhp_zips <- c(sfmhp$RegionName)
sfmhp_zips
```

4. Google API to convert Zip Code to Lat-Long

```{r}
library(RCurl)
library(RJSONIO)
library(plyr)
```

```{r}
url <- function(address, return.call = "json") {
 root <- "http://maps.google.com/maps/api/geocode/"
 u <- paste(root, return.call, "?address=", address, sep = "")
 return(URLencode(u))
}
```

```{r}
geoCode <- function(address,verbose=FALSE) {
 if(verbose) cat(address,"\n")
 u <- url(address)
 print (u)
 doc <- getURL(u)
 x <- fromJSON(doc,simplify = FALSE)
 if(x$status=="OK") {
 lat <- x$results[[1]]$geometry$location$lat
 lng <- x$results[[1]]$geometry$location$lng
 location_type <- x$results[[1]]$geometry$location_type
 formatted_address <- x$results[[1]]$formatted_address
 zip_code <- address
 return(c(zip_code, lat, lng, location_type, formatted_address))
 Sys.sleep(0.5)
 } else {
 return(c(NA,NA,NA,NA,NA))
 }
}
```

```{r}
##Test with a single address

# Use plyr to ggeocoding for a vector
locations_df <- data.frame(lat=character(),
                 lon=character(), 
                 location_type=character(),
                 formatted=character(),
                 stringsAsFactors=FALSE) 
address <- c(sfmhp_zips)
locations <- ldply(address, function(x) geoCode(x))
locations_df = rbind(locations_df,locations)
names(locations_df) <- c("zip_code", "lat", "lon", "location_type", "formatted")
head(locations_df)

smallset <- locations_df$zip_code[!is.na(locations_df$zip_code)]
smallset

bigset <- c(sfmhp$RegionName)
bigset

setdiff(bigset, smallset)

# Writing Zip_Codes and Lat Long data to a file
write.table(locations_df, file = "/Users/gsista/Desktop/Springboard/DataScience/GoogleAPILocation.tsv", sep = "\t",
            row.names = TRUE, col.names = NA)

```

5. Google API to find the distance in miles between two Lat-Longs

```{r}
url_latlong <- function(origins, destinations) {
 root <- "https://maps.googleapis.com/maps/api/distancematrix/json?units=imperial"
 u <- paste(root, "&origins=", origins, "&destinations=", destinations, sep = "")
 return(URLencode(u))
}
```

```{r}
findapproxmatch <- function(origins, destinations,threshold, verbose=FALSE) {
 if(verbose) cat(origins, destinations, threshold,"\n")
 u <- url_latlong(origins, destinations, threshold)
 print (u)
 doc <- getURL(u)
 x <- fromJSON(doc,simplify = FALSE)
 print(x)
}

if(x$status=="OK") {
  origins <- $origin_addresses[[1]]
  destinations <- 
#  return(c(zip_code, lat, lng, location_type, formatted_address))
#  Sys.sleep(0.5)
#  } else {
#  return(c(NA,NA,NA,NA,NA))
#  }
}
```

```{r}
threshold = "10"
origins = "37.7929789,-122.4212424"
destinations = "37.7225491,-122.4410618"
var <- geoCode(origins, destinations)
var$rows[[1]]$elements[[1]]$distance$text[1]
```

