---
title: "Springboard Capstone Project"
author: Gautham Sista
output: html_notebook
---

This notebook shows the analysis performed as part of my Springboard Capstone Project for the Introduction to Data Science course. 

1a. Reading in the San Francisco Crime dataset

```{r}
sf <- read.csv("/Users/gsista/Desktop/Springboard/DataScience/Data/sf_crime.csv")
sf$Dates <- as.POSIXct(sf$Dates, format="%Y-%m-%d %H:%M:%S")
str(sf)
```

1b. Filtering the San Francisco dataset to crimes that happened on and after January 1st, 2013
```{r}
sf1 <- subset(sf, Dates >= "2013-01-01 00:00:00")
sf1$LatLong <- NA
sf1$LatLong <- paste(sf1$Y, sf1$X, sep = ", ")
unique(sf1$LatLong)
```

```{r}
sf$DayOfWeek <- factor(sf$DayOfWeek, levels=c('Monday', 'Tuesday', 'Wednesday', 'Thursday', 'Friday', 'Saturday', 'Sunday'))
library(ggplot2)
ggplot(sf, aes(x = DayOfWeek)) +
  geom_bar() +
  geom_text(stat='count',aes(label=..count..),vjust=-1)
```

```{r}
sf$DayOfWeek <- factor(sf$DayOfWeek, levels=c('Monday', 'Tuesday', 'Wednesday', 'Thursday', 'Friday', 'Saturday', 'Sunday'))
ggplot(sf, aes(x = Category, fill = DayOfWeek)) +
  geom_bar() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```


2. Reading in the Zipcodes Latitude and Longitude dataset

```{r}
zip <- read.csv("/Users/gsista/Desktop/Springboard/DataScience/Springboard-Capstone-Project/Zipcodes_LatLong.txt", sep = "")
str(zip)
zip
```

3. Reading in Median House Prices by zipcode dataset

```{r}
mhp <- read.csv("/Users/gsista/Desktop/Springboard/DataScience/Springboard-Capstone-Project/Zip_MedianListingPrice_AllHomes.csv")
str(mhp)
mhp
```

4. 